---
title: "CaseStudy2"
author: "Anthony Burton-Cordova"
date: "2022-11-19"
output: html_document
---

```{r, echo=FALSE}
###Packages
library(XML)
library(dplyr)
library(tidyr)
library(stringi)
library(rvest) 
library(ggplot2)
library(RCurl)
library(class)
library(caret)
library(e1071)
library(stringr)
library(naniar)
library(rmarkdown)
library(readxl)
```
  
```{r,echo=FALSE, message=FALSE,warning=FALSE,results="hide"}
###Import Employee Data

employeeData = read.csv(file.choose(), sep = ",")
head(employeeData)
```

## Cleaning Dataset
In this section , I will be checking for nulls and encoding columns with characters variables to numeric variables. 
```{r,echo=FALSE, message=FALSE,warning=FALSE,results="hide"}
### Check for NA's
sum(is.na(employeeData))

# Make overtime column binary 
employeeData$cleanOverTime = ifelse(employeeData$OverTime=="Yes",1,0)
employeeData$cleanAttrition = ifelse(employeeData$Attrition=="Yes",1,0)

# recode JobRole column to numeric values
recode_values = function(x) {
  relist(match(unlist(x), unique(unlist(x))), x)
}

employeeData$cleanDepartment = recode_values(employeeData$Department)
employeeData$cleanEducation = recode_values(employeeData$Education)
employeeData$cleanEducField = recode_values(employeeData$EducationField)
employeeData$cleanGender = recode_values(employeeData$Gender)
employeeData$cleanJobRole = recode_values(employeeData$JobRole)
employeeData$cleanMarStatus = recode_values(employeeData$MaritalStatus)
employeeData$cleanBusinessTravel = recode_values(employeeData$BusinessTravel)

#employeeData = employeeData
#employeeData$MonthlyIncome=scale(employeeData$MonthlyIncome)
#employeeData$MonthlyRate=scale(employeeData$MonthlyRate)
#employeeData$JobSatisfaction=scale(employeeData$JobSatisfaction)
#employeeData$NOverTime=scale(employeeData$NOverTime)
#employeeData$Age=scale(employeeData$Age)
#employeeData$EnvironmentSatisfaction=scale(employeeData$EnvironmentSatisfaction)
#employeeData$DailyRate=scale(employeeData$DailyRate)
#employeeData$JobLevel=scale(employeeData$JobLevel)
```

## Attrition Exploratory Data Analysis(EDA)
In this section, we will perform Exploratory Data Analysis(EDA) on various variables that we believe will lead to attrition.
From our analysis, we find that Job Role seems to be one factor that leads to attrition since Sales Representative are more likely to quit. We find that the attrition rate is greater than 40%.
Monthly Income also seems to fit since as we move up in income, we find that the attrition rate is getting lower, almost nonexistent after the halfway mark of $10,000.

```{r,echo=FALSE, message=FALSE,warning=FALSE,results="hide"}
### Exploratory Data Analysis

employeeData %>% ggplot(aes(x=Attrition,fill=Attrition)) + 
  geom_bar()+
  ggtitle("Attrition Count") +
  xlab("Attrition")+ylab("Count")

### Compare Incomes
employeeData %>% group_by(Attrition) %>% summarise(compareincomes=mean(MonthlyIncome))

### Percentage Compares for Job Role
ggplot(employeeData, aes(x = JobRole, fill = Attrition)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent)+ 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))


### Attrition Vs. Job Satisfaction
employeeData %>% 
  ggplot(aes(x=JobSatisfaction,fill=Attrition))+
  geom_bar()+
  ggtitle("Attrition Vs. Job Satisfaction") 

### Attrition Vs. MonthlyIncome
employeeData %>% ggplot(aes(x=MonthlyIncome,fill=Attrition))+
  geom_histogram()+
  ggtitle("Attrition Vs. MonthlyIncome") 

### Attrition Vs. DistanceFromHome
employeeData %>% 
  ggplot(aes(x=DistanceFromHome,fill=Attrition))+
  geom_histogram()+
  ggtitle("Attrition Vs. DistanceFromHome") 

### Attrition Vs. MonthlyRate
employeeData %>% ggplot(aes(x=MonthlyRate,fill=Attrition))+geom_histogram()+ggtitle("Attrition Vs. MonthlyRate") 

### Attrition Vs. OverTime
employeeData %>% 
  ggplot(aes(x=OverTime,fill=Attrition))+
  geom_bar(position="fill")+ggtitle("Attrition Vs. Overtime")+
  scale_y_continuous(labels = scales::percent)

### Attrition Vs. years Since last Promotion
employeeData %>% 
  ggplot(aes(x=YearsSinceLastPromotion,fill=Attrition))+
  geom_bar(position="fill")+ggtitle("Attrition Vs. Years Since Last Promotion") +  
  scale_y_continuous(labels = scales::percent)

### Attrition Vs. Salary Hike - NO
employeeData %>% 
  ggplot(aes(x=PercentSalaryHike,fill=Attrition))+
  geom_histogram(position="fill")+
  ggtitle("Attrition Vs. Percent Salary Hike") 

### Attrition Vs. Age
employeeData %>% 
  ggplot(aes(x=Age,fill=Attrition))+
  geom_histogram()+
  ggtitle("Attrition Vs. Age")

### Attrition vs Marital Status
employeeData %>% 
  ggplot(aes(x=MaritalStatus,fill=Attrition))+
  geom_bar(position="fill")+
  ggtitle("Attrition Vs. Marital Status")

### Attrition Vs. PercentSalaryHike
employeeData %>% 
  ggplot(aes(x=PercentSalaryHike,fill=Attrition))+
  geom_histogram()+ggtitle("Attrition Vs. PercentSalaryHike")+
  scale_y_continuous(labels = scales::percent)

### Attrition Vs. WorkLifeBalance
employeeData %>% 
  ggplot(aes(x=WorkLifeBalance,fill=Attrition))+
  geom_histogram()+ggtitle("Attrition Vs. WorkLifeBalance")+
  scale_y_continuous(labels = scales::percent)

### Attrition Vs. StockOptionLevel
employeeData %>% 
  ggplot(aes(x=StockOptionLevel,fill=Attrition))+
  geom_histogram()+ggtitle("Attrition Vs. StockOptionLevel")+
  scale_y_continuous(labels = scales::percent)

### Attrition Vs. TrainingTimesLastYear
employeeData %>% 
  ggplot(aes(x=TrainingTimesLastYear,fill=Attrition))+
  geom_histogram()+
  ggtitle("Attrition Vs. TrainingTimesLastYear")+
  scale_y_continuous(labels = scales::percent)

### Attrition Vs. PerformanceRating
employeeData %>% 
  ggplot(aes(x=PerformanceRating,fill=Attrition))+
  geom_histogram()+ggtitle("Attrition Vs. PerformanceRating")+
  scale_y_continuous(labels = scales::percent)

### Attrition Vs. RelationshipSatisfaction
employeeData %>% 
  ggplot(aes(x=RelationshipSatisfaction,fill=Attrition))+
  geom_histogram()+ggtitle("Attrition Vs. RelationshipSatisfaction")+
  scale_y_continuous(labels = scales::percent)

### Attrition Vs. BusinessTravel
employeeData %>% 
  ggplot(aes(x=BusinessTravel,fill=Attrition))+
  geom_bar()+ggtitle("Attrition Vs. BusinessTravel")+
  scale_y_continuous(labels = scales::percent)
```

## KNN Model
```{r,echo=FALSE, message=FALSE,warning=FALSE,results="hide"}

### KNN Model

## Classify
employeeData2 = employeeData

#standardize data
employeeData2$MonthlyIncome=scale(employeeData2$MonthlyIncome)
employeeData2$MonthlyRate=scale(employeeData2$MonthlyRate)
employeeData2$JobSatisfaction=scale(employeeData2$JobSatisfaction)
employeeData2$cleanOverTime=scale(employeeData2$cleanOverTime)
employeeData2$Age=scale(employeeData2$Age)
employeeData2$EnvironmentSatisfaction=scale(employeeData2$EnvironmentSatisfaction)
employeeData2$DailyRate=scale(employeeData2$DailyRate)
employeeData2$JobLevel=scale(employeeData2$JobLevel)

model = employeeData2 %>% select(c("MonthlyIncome","MonthlyRate","Age","cleanOverTime",
                                   "JobSatisfaction","PerformanceRating","JobLevel",
                                   "YearsSinceLastPromotion","YearsInCurrentRole" ,"Attrition"))
head(model)

## ML
#Set Seed
iterations = 200
numks = 20
splitPerc = .70
masterAcc = matrix(nrow = iterations, ncol = numks)
for(j in 1:iterations)
{
  trainIndices = sample(1:dim(model)[1],round(splitPerc * dim(model)[1]))
  train = model[trainIndices,]
  test = model[-trainIndices,]
  for(i in 1:numks)
  {
    classifications = knn(train[,c(1,4,7)],test[,c(1,4,7)],train$Attrition, prob = TRUE, k = i)
    table(classifications,test$Attrition)
    CM = confusionMatrix(table(classifications,test$Attrition))
    masterAcc[j,i] = CM$overall[1]
  }
  
}
MeanAcc = colMeans(masterAcc)
plot(seq(1,numks,1),MeanAcc, type = "l")
which.max(MeanAcc)
max(MeanAcc)
```

```{r,echo=FALSE, message=FALSE,warning=FALSE,results="hide"}
classifications = knn(train[,c(1,4,7)],test[,c(1,4,7)],train$Attrition, prob = TRUE, k = 15)
table(classifications,test$Attrition)
CM = confusionMatrix(table(classifications,test$Attrition))
CM
```

```{r,echo=FALSE, message=FALSE,warning=FALSE,results="hide"}
set.seed(13)
naive_data=employeeData2

model2 = naive_data %>% select(c("MonthlyIncome","JobRole","Age","cleanOverTime",
                                 "EnvironmentSatisfaction","DailyRate","JobLevel",
                                 "TotalWorkingYears","JobInvolvement","JobSatisfaction",
                                 "Attrition"))
model2$Attrition = as.factor(model2$Attrition)

trainIndices = sample(1:dim(model2)[1],round(.70 * dim(model2)[1]))
train = model2[trainIndices,]
test = model2[-trainIndices,]

classifier1 = naiveBayes(model2[,c(1,4,7)],model2$Attrition)

pred = predict(classifier1,newdata=test)
CM = confusionMatrix(table(test$Attrition,pred))

CM
```


```{r,echo=FALSE, message=FALSE,warning=FALSE,results="hide"}
lm_data = employeeData2
lm_data$Attrition = as.factor(lm_data$Attrition)
levels(lm_data$Attrition) = list(YES = "Yes", NO = "No")

#num_Attrition = ifelse(lm_data$Attrition=="Yes",1,0)
#lm_data$cleanAttrition = num_Attrition
fit = lm(cleanAttrition~MonthlyIncome+cleanOverTime+JobLevel,data = lm_data)
summary(fit)
```

```{r,echo=FALSE, message=FALSE,warning=FALSE,results="hide"}
employee_nosalary = read.csv(file.choose())

print(employeeData)
```

## Testing Correlation and Choosing Variables For Linear Regression Model to Predict Monthly Income(Salary)
```{r,echo=FALSE, message=FALSE,warning=FALSE,results="hide"}
cor(employeeData$MonthlyIncome,employeeData[,c(2,5,7,8,12,14,15,16,18,20,
                                                21,25,26,27,29,30,31,32,33,34,
                                                35,36,37,38:45)])
lm_salarydf = employeeData[,c(2,5,7,8,12,14,15,16,18, 20,
                                                21,25,26,27,29,30, 31,32,33,34,
                                                35,36,37,38:45)]
lmsalary_model = lm(MonthlyIncome~.,
                    data = lm_salarydf)
summary(lmsalary_model)
```

```{r,echo=FALSE, message=FALSE,warning=FALSE,results="hide"}
lm_salarydf = employeeData[,c(7, 16, 20, 30, 36, 37, 39, 41, 43, 45)]

lmsalary_model = lm(MonthlyIncome~JobLevel+TotalWorkingYears+YearsWithCurrManager
                    +DistanceFromHome+ cleanJobRole + cleanDepartment 
                    + cleanBusinessTravel,
                    data = lm_salarydf)
summary(lmsalary_model)

RSS = c(crossprod(lmsalary_model$residuals))
MSE = RSS / length(lmsalary_model$residuals)
RMSE = sqrt(MSE)
sig2 = RSS / lmsalary_model$df.residual

RMSE
```

## Predicting Salary
```{r,echo=FALSE, message=FALSE,warning=FALSE,results="hide"}
# recode catergorical values to numeric values for model 
employee_nosalary$cleanDepartment = recode_values(employee_nosalary$Department)
employee_nosalary$cleanJobRole = recode_values(employee_nosalary$JobRole)
employee_nosalary$cleanBusinessTravel = recode_values(employee_nosalary$BusinessTravel)

salary_pred = predict(lmsalary_model, employee_nosalary)
salary_pred
```

## Adding Column to No Salary Dataset and Exporting as .csv File
```{r,echo=FALSE, message=FALSE,warning=FALSE,results="hide"}
salary_pred = unlist(salary_pred)

employee_nosalary$SalaryPred =salary_pred
head(employee_nosalary)
```


## Adding Column to No Salary Dataset and Exporting as .csv File
```{r,echo=FALSE, message=FALSE,warning=FALSE,results="hide"}
### Bring in the No Attrition Data Set
employee_noattrition = read.csv(file.choose())

print(employee_noattrition)
```
```{r,echo=FALSE, message=FALSE,warning=FALSE,results="hide"}
###Using a Naive Bayes model to predict attrition

naive_data=employeeData

model2 = naive_data
model2$Attrition = as.factor(model2$Attrition)

trainIndices = sample(1:dim(model2)[1],round(.70 * dim(model2)[1]))
train = model2[trainIndices,]
test = model2[-trainIndices,]

classifier1 = naiveBayes(model2[,-c(1,5,9,10,11,14,21,23,24,28,37:45)],model2$Attrition)

pred = predict(classifier1,newdata=test)
CM = confusionMatrix(table(test$Attrition,pred))

CM

```